The ChEMBL database as Linked Open Data

### 2012-04-16 article.tex
Filling in Data Statistics and Validation section

John/Anna: common statistics with SPARQL

>>>>

Each release of the ChEMBL database is accompanied by set of release counts, which help 
provide a concise content overview of the release in question. The ChEMBL 13 release counts
are as follows:

\begin{itemize}
  \item 8,845 targets
  \item 1,143,682 distinct molecules
  \item 6,933,068 activities
  \item 617,681 assays
  \item 44,682 documents
\end{itemize}

These counts are important as firstly they show continued growth in ChEMBL data, as seen 
in all previous releases. Secondly, these counts can help validate any transformations, 
which have been applied to the relational data model traditionally used to store the ChEMBL 
data. The following SPARQL queries attempt to provide a high level validation of the 
ChEMBL-RDF by regenerating the ChEMBL 13 release statistics.

The following SPARQL query returns 8,845 ChEMBL targets:

\begin{tiny}
\begin{verbatim}
select count(*) WHERE {
  ?s ?p chembl:Target .
}
\end{verbatim}
\end{tiny}

The following SPARQL query returns 1,143,682 ChEMBL molecules:

NOTE: Full set of ChEMBL molecules have been created by querying molecule_dictionary table
NOTE: I have added extra triple for each molecule: <http://data.kasabi.com/dataset/chembl-rdf/13/molecule/m1> <http://www.w3.org/1999/02/22-rdf-syntax-ns#type> <http://rdf.farmbio.uu.se/chembl/onto/#Molecule> .
NOTE: Modified molecule compound structure can be queried here: https://wwwdev.ebi.ac.uk/chembl/sparql

\begin{tiny}
\begin{verbatim}
select count(*) WHERE {
  ?s ?p chembl:Molecule .
}
\end{verbatim}
\end{tiny}

The following SPARQL query returns 617,681 ChEMBL assays:

\begin{tiny}
\begin{verbatim}
select count(*) WHERE {
  ?s ?p chembl:Assay .
}
\end{verbatim}
\end{tiny}

The following SPARQL query returns the 6,933,068 ChEMBL activities:

\begin{tiny}
\begin{verbatim}
select count(*) WHERE {
  ?s ?p chembl:Activity .
}
\end{verbatim}
\end{tiny}

The following SPARQL query returns the 44,681 ChEMBL documents. This is 1 less than appears 
in ChEMBL 13 release notes, as it does not include doc_id -1, which refers to the 
unpublished dataset.

\begin{tiny}
\begin{verbatim}
select count(*) WHERE {
  ?s ?p bibo:Article .
}
\end{verbatim}
\end{tiny}

Recreation of the ChEMBL release statistics, by querying the ChEMBL-RDF with SPARQL 
provides end users with high degree of confidence that the ChEMBL-RDF contains same 
core content as source relational database.
